
\chapter{Literature Review}\label{chap:litreview}

\section{Review Scope and Questions}
This chapter reviews the literature most relevant to the thesis problem of securing conversational audio systems against spoofing, impersonation, and provenance ambiguity. The review spans three tightly connected areas: (i) attacks and defensive requirements in voice interfaces; (ii) audio watermarking for synthetic-speech provenance; and (iii) deepfake and replay detection for user-originated audio. The implemented prototype of the present thesis belongs to the second area, namely assistant-output watermarking, but the third area remains important because it motivates the broader middleware architecture discussed throughout the dissertation.

\section{Review Method (Concise Summary)}
A structured narrative review was used instead of a full systematic literature review. The process followed four simple steps: first, we defined a taxonomy aligned with the bidirectional nature of conversational systems; second, we searched major digital libraries, arXiv, conference proceedings, and official technical webpages using combinations of the keywords \emph{voice assistant security}, \emph{audio watermarking}, \emph{speech provenance}, \emph{deepfake speech}, \emph{replay detection}, and \emph{ASVspoof}; third, we prioritized peer-reviewed work and primary technical sources when available; and fourth, we synthesized the findings by theme. The goal of the review is therefore not exhaustive coverage of all audio forensics, but rather a technically grounded synthesis that explains why the present prototype was designed the way it was.

\section{Taxonomy Used in This Review}
The reviewed work is organized into four themes:
\begin{enumerate}[leftmargin=1.25em]
  \item \textbf{Threats and attacks in voice interfaces} (replay, injected commands, TTS/VC impersonation, deepfakes);
  \item \textbf{Audio watermarking for speech provenance} (zero-bit and payload-bearing watermarking, robustness, imperceptibility, localization, deployment constraints);
  \item \textbf{Deepfake and replay detection} (artifact-based detection, generalization, calibration, channel mismatch);
  \item \textbf{Hybrid and integrated defenses} (systems that combine provenance, attribution, and detector signals).
\end{enumerate}
This taxonomy is useful because the thesis itself sits at the boundary between provenance and detection: the current implementation realizes only the provenance branch, while the full long-term architecture requires both.

\section{Voice Assistant Security and AI-Generated Audio Threats}
Voice assistants operate through an unusually exposed interface: they continuously accept acoustic input and often produce trusted speech output intended to guide user actions. Prior work has long documented replay attacks and injected commands in such systems\,[15]. Recent advances in voice cloning and speech generation have substantially amplified the threat surface. With modern TTS and voice conversion, attackers can synthesize speech that is close enough to a target speaker to fool listeners and, in some cases, deployed verification systems\,[3][16][17][18].

\paragraph{Two attack directions.}
For conversational systems, the threat model is naturally bidirectional.
\begin{itemize}[leftmargin=1.25em]
  \item \textbf{User $\rightarrow$ assistant.} An attacker attempts to impersonate a user or inject commands by replaying prior recordings or synthesizing speech with TTS/VC.
  \item \textbf{Assistant $\rightarrow$ user.} A malicious or tampered audio stream pretends to originate from the assistant, for example in phishing, telephony impersonation, or man-in-the-middle scenarios.
\end{itemize}
The first direction motivates user-side deepfake and replay detection. The second direction motivates watermarking of assistant speech, because provenance can then be checked by software instead of by human listening alone.

\section{Audio Watermarking for Speech Provenance}
Audio watermarking embeds a hidden but machine-detectable pattern into an audio signal. In the context of synthetic speech, watermarking is attractive because it transforms provenance checking into a direct verification problem: rather than infer whether a waveform \emph{looks synthetic}, a verifier can test whether it contains the expected mark\,[21][22]. This is especially relevant for assistant speech, where the generator is known and can embed a watermark proactively.

\paragraph{Representative design axes.}
The recent literature reveals several recurring trade-offs.
\begin{itemize}[leftmargin=1.25em]
  \item \textbf{Zero-bit vs. payload-bearing watermarking.} Zero-bit designs optimize presence/absence detection and often maximize robustness, whereas payload-bearing systems can encode identifiers, model versions, or other metadata at the cost of a harder decoding problem\,[35][36][37].
  \item \textbf{Localization granularity.} Some systems return only a global confidence score, whereas others explicitly localize where the watermark is present in time\,[14][25][26][27]. Localization matters whenever audio can be cropped, edited, or consumed incrementally.
  \item \textbf{Robustness vs. perceptual transparency.} Watermarks should survive compression, noise, filtering, and editing, but stronger residuals can damage speech quality. Much recent work explicitly trades off robustness against PESQ/MOS-like fidelity objectives\,[22][28][32][33][34].
  \item \textbf{Deployment target.} Not all systems solve the same problem. Some watermark generated speech directly; others protect source recordings so that subsequent voice cloning transfers the watermark into the fake speech\,[23][24].
\end{itemize}

\paragraph{Representative approaches.}
AudioSeal is one of the clearest recent references for localized speech watermarking: it jointly trains a generator and a fast detector, uses a localization loss, and emphasizes robustness together with perceptual masking\,[14][25][26][27][56]. TRUE instead places particularly strong emphasis on fidelity while remaining robust, and reports very high objective speech quality for speech and singing voice\,[32][33][34]. AudioMarkNet addresses a related but different problem: it watermarks source recordings so that later voice-cloning systems inherit the watermark during speaker adaptation, thereby turning fake-speech detection into a watermark retrieval problem\,[23][24]. Proprietary industrial systems such as SynthID demonstrate that watermarking is now viewed not only as an academic topic but also as a deployable transparency mechanism for AI-generated audio\,[55]. Finally, FakeMark extends the discussion toward attribution, showing that watermark-like cues can be coupled with generator-specific artifacts so that the detector is not limited to binary presence detection\,[10][50][51][52][53][54].

\section{Deepfake and Replay Audio Detection}
The literature on user-side audio integrity has developed along a partly separate path. Classical systems relied on cepstral features such as MFCC and CQCC with conventional classifiers, while modern systems increasingly use convolutional, transformer-based, or self-supervised speech encoders\,[38][39][40][5]. Across this literature, the main unresolved issue is not in-domain accuracy but \emph{generalization}. Many detectors perform well on known attacks but degrade on unseen generators, codecs, microphones, reverberation conditions, or replay channels\,[41][42][6][8][43].

\paragraph{Why channel realism matters.}
This observation is directly relevant to the current thesis. If a defense is evaluated only on clean offline signals, the reported numbers may be optimistic for deployment. Real systems involve compression, transmission, room acoustics, and device-specific colorations. Prior work therefore stresses augmentation, calibration, and realistic channel simulation when assessing deepfake detection and provenance mechanisms\,[44][45][13]. This point strongly influenced the methodology of the present thesis: even though the implemented benchmark remains modest, it deliberately includes codecs, filtering, and resampling-based perturbations instead of reporting clean-only numbers.

\section{Hybrid and Integrated Defenses}
The strongest systems-level conclusion from the literature is that no single signal is likely to remain sufficient forever. Pure artifact-based detection may fail under distribution shift, while pure watermarking may be weakened by severe channel distortion or targeted removal. Hybrid designs therefore attempt to combine complementary cues. FakeMark is especially important here because it makes this idea explicit: the system couples watermark-like signatures with intrinsic generation artifacts so that attribution can still succeed even if one cue weakens\,[50][51][52][53][54].

This layered perspective motivates the broader architecture discussed in the present thesis. The implemented prototype should therefore be interpreted as the \emph{assistant-output provenance layer} of a future bilateral middleware rather than as a complete end-to-end anti-spoofing solution.

\section{Synthesis: What the Literature Implies for Our Design}
Four design lessons follow from the reviewed work.
\begin{itemize}[leftmargin=1.25em]
  \item \textbf{Provenance should be proactive.} If the assistant controls its own TTS pipeline, embedding the watermark at generation time is preferable to relying only on downstream artifact detection\,[14][25][26][27].
  \item \textbf{Localization is operationally important.} Edited, cropped, or streamed audio is easier to verify when the detector predicts where the watermark is present instead of outputting only a global score\,[14][25][26][27].
  \item \textbf{Perceptual quality must be measured explicitly.} Strong detection performance is not enough if the watermark remains audible. Recent systems such as TRUE make fidelity a first-class objective rather than an afterthought\,[32][33][34].
  \item \textbf{Channel diversity must be part of the benchmark.} Codec and transmission realism are essential for honest claims about robustness\,[44][45][13].
\end{itemize}
These lessons explain the current prototype's emphasis on localized detection and prefix-aware verification, while also clarifying two of its current weaknesses: perceptual evaluation is still too limited, and channel diversity is narrower than in a fully deployment-oriented benchmark.

\subsection{Our design vs. prior art}
The present thesis differs from prior work in several specific ways.
\begin{itemize}[leftmargin=1.25em]
  \item \textbf{The implemented scope is intentionally narrow.} Unlike fully hybrid proposals, the current prototype implements only the assistant-output watermarking branch. The user-side detector remains a future integration target rather than a claim of current functionality.
  \item \textbf{Prefix-aware verification is a central objective.} Relative to prior localized watermarking systems, the present work places unusual emphasis on whether the watermark can already be verified from very short prefixes (0.25/0.5/1.0/2.0~s), which is directly relevant to interactive assistants.
  \item \textbf{The system uses a small payload instead of strict zero-bit detection.} The implemented checkpoint embeds a 4-bit payload and decodes it with masked statistics pooling over the predicted watermark region.
  \item \textbf{Functionality was prioritized before perceptual refinement.} In contrast to work such as TRUE, the strongest current checkpoint was optimized first for detection, localization, and low-latency decoding, while explicit STFT/$\ell_2$ quality penalties were deferred to a later engineering stage.
  \item \textbf{The benchmark claims are correspondingly conservative.} The reported evidence is restricted to 16~kHz speech on LibriSpeech plus a concrete black-box attack suite, rather than broad claims about arbitrary acoustics, languages, or end-to-end telephony.
\end{itemize}

\subsection{Comparison with representative recent systems}
Table~\ref{tab:related_systems_compare} positions the current prototype against five representative systems frequently discussed in recent work: AudioSeal\,[14][25][26][27][56][73], TRUE\,[32][33][34], AudioMarkNet\,[23][24], SynthID\,[55], and FakeMark\,[10][50][51][52][53][54]. Two cautions are important. First, the systems do not all solve exactly the same task. Second, for proprietary or partially disclosed systems, some fields are not publicly specified and are therefore marked as such. Even with these caveats, the comparison is useful because it makes explicit that the novelty of the present prototype lies less in raw payload size and more in the combination of localized detection, low false-positive operation, and prefix-aware verification.

\begin{table}[H]
\centering
\scriptsize
\setlength{\tabcolsep}{3pt}
\caption{Comparison of the current prototype with representative recent audio provenance and attribution systems.}
\label{tab:related_systems_compare}
\begin{tabular}{p{1.8cm}p{1.5cm}p{1.6cm}p{1.6cm}p{2.1cm}p{1.8cm}p{1.7cm}p{2.5cm}}
\toprule
\textbf{System} & \textbf{Payload} & \textbf{Localization} & \textbf{Prefix-aware} & \textbf{Robustness target} & \textbf{Perceptual focus} & \textbf{Domain} & \textbf{Evaluation attacks} \\
\midrule
This work & 4-bit payload & frame-level presence map + masked pooling & yes, explicit 0.25--2.0~s probes & assistant-speech provenance under common black-box edits & secondary in current checkpoint; quality losses currently zero & 16~kHz speech (LibriSpeech) & noise, speed/resample, low-pass, crop-pad, bit-depth, MP3, AAC; code also prepared for phone-band/Opus \\
\addlinespace[0.15em]
AudioSeal & optional 16-bit secret message on top of presence detection & strong localized detection (sample/frame-level) & real-time/streaming friendly, but not centered on prefix payload decoding & robust detection after cropping, re-encoding, and noise & strong, explicit perceptual masking objective & speech / AI-generated speech & cropping/editing, re-encoding, noise, common manipulations \\
\addlinespace[0.15em]
TRUE & payload-bearing watermark & not primarily presented as localized detector & no explicit prefix-verification emphasis & robustness with very strong fidelity preservation & very strong; high PESQ emphasized & speech and singing voice & broad common audio distortions in paper benchmark \\
\addlinespace[0.15em]
AudioMarkNet & registered watermark identity rather than small user payload & coarse 1~s section-wise retrieval & coarse only; 1~s sections/UFL rather than ASR-aligned prefixes & transfer through speaker adaptation and resistance to non-adaptive/adaptive attacks & moderate; quality reported but not the primary novelty & original speech $\rightarrow$ cloned fake speech & noise, resampling, adaptive removal, TTS transfer scenarios \\
\addlinespace[0.15em]
SynthID & not publicly disclosed for audio & not publicly disclosed & not publicly disclosed & robustness to common user edits in product setting & product-level inaudibility is a central claim & Google-generated audio products & public claims cover noise, MP3 compression, and speed changes \\
\addlinespace[0.15em]
FakeMark & system-correlated signature rather than fixed bitstring & localization is not the primary goal & no & source attribution under domain shift and watermark removal pressure & balanced, but attribution robustness is primary & deepfake speech attribution & codec compression, removal attacks, cross-dataset evaluation \\
\bottomrule
\end{tabular}
\end{table}

The comparison highlights a useful positioning statement for the thesis. AudioSeal is the closest reference point in terms of localized speech watermarking. TRUE is the closest reference point in terms of fidelity-oriented optimization. AudioMarkNet and FakeMark widen the design space by showing that watermarking can serve not only binary provenance detection but also transfer-based detection and source attribution. SynthID, in turn, demonstrates that industrial deployment places a premium on practicality and user-facing verification, even when many technical details remain undisclosed. Against this background, the present prototype is best described as a \emph{research-oriented localized and prefix-aware speech watermark}, rather than as the final word on either fidelity or product-scale deployment.

\section{Gaps and Research Opportunities}
The literature still leaves several open problems directly relevant to the thesis.
\begin{itemize}[leftmargin=1.25em]
  \item \textbf{Targeted watermark removal vs. speech quality.} The trade-off between stealth, robustness, and resistance to adaptive removal remains under-characterized for speech.
  \item \textbf{Cross-channel robustness.} More work is needed on telephony, realistic VoIP paths, room acoustics, and neural codec pipelines.
  \item \textbf{Short-prefix operating points.} Most work reports full-utterance accuracy, whereas interactive systems need trustworthy early decisions under partial context.
  \item \textbf{Integrated bilateral middleware.} Few systems jointly quantify assistant-output provenance, user-input integrity, and conversational latency within one end-to-end benchmark.
\end{itemize}
These gaps motivate the present thesis directly: the current implementation tackles localized, low-latency assistant-output watermarking first, while leaving fidelity-oriented refinement, broader channel realism, and bilateral integration to the next stage of work.

\section{Threats to the Validity of This Review}
Three limitations should be acknowledged. First, the review is intentionally centered on voice-assistant security, so broader audio forensics is covered only selectively. Second, because the review is structured and narrative rather than fully systematic, some selection bias remains possible despite the focus on primary sources. Third, several industrial systems, most notably SynthID, disclose only high-level properties publicly; any comparison involving them must therefore distinguish between what is explicitly documented and what is simply not public.
