
\chapter{Introduction}

\section{Background and Motivation}
Virtual voice assistants have become ubiquitous since the introduction of Apple’s Siri in 2011, offering convenient hands-free interaction through natural speech\,[1]. As these conversational agents handle an expanding range of sensitive tasks (from smart-home control to banking), ensuring their safety and security has grown paramount. Unfortunately, voice-based systems are inherently vulnerable to spoofing and impersonation attacks---malicious actors can exploit the interface by injecting falsified audio commands or masquerading as a trusted speaker. Recent advances in AI-driven voice cloning (deepfakes) have enabled attackers to generate synthetic speech that is nearly indistinguishable from a target speaker, leading to real-world incidents. For example, in 2019 fraudsters used AI to mimic a CEO’s voice and trick an employee into a \$243{,}000 bank transfer\,[2]. Likewise, researchers have demonstrated that even state-of-the-art voice authentication systems (e.g., WeChat Voiceprint, Microsoft Azure, Amazon Alexa’s voice ID) can be bypassed by cloned voices, highlighting the severity of this threat\,[3].

\section{Problem Statement and Research Gap}
Existing countermeasures typically fall into two categories: (i) user authentication and liveness checks, and (ii) deepfake detection algorithms. The former (e.g., voice biometrics, challenge--response) can be intrusive and are known to be defeasible via replayed or synthesized responses. The latter (artifact-based detection) has seen rapid progress\,[4][5] but remains brittle under distribution shift: detectors often struggle with unseen attacks, codecs, recording chains, or acoustic environments, causing error rates to climb markedly outside controlled conditions\,[6][7][8]. In interactive assistants, this mismatch undermines safety at the exact moment of deployment.

The broader system vision motivating this thesis is a layered defense that combines (a) a lightweight provenance check for assistant-originated audio and (b) a detector for forged user-originated audio. However, the \emph{implemented and empirically evaluated} work in the present thesis is narrower: it focuses on \textbf{assistant-output watermarking with localized, low-latency detection on LibriSpeech}. The complementary user-side deepfake/replay detector is retained as a \textbf{conceptual extension and future work}, not as an implemented experimental branch.

\section{Objectives and Research Questions}
The long-term objective is a bilateral safety layer for conversational agents. In the current thesis, however, the implemented prototype operationalizes only the \emph{assistant-output watermarking branch}. Accordingly, the research questions are narrowed to match the empirical scope that is actually covered by the code base and reported experiments.

\begin{itemize}[leftmargin=1.25em]
  \item \textbf{RQ1 (Perceptual impact, partially addressed):} Can a TTS watermark be embedded in assistant speech while preserving usable speech quality, and what remains to reach true imperceptibility in the current prototype?
  \item \textbf{RQ2 (Robustness on the implemented benchmark):} On \textbf{LibriSpeech} 16~kHz speech, does the watermark remain detectable and decodable after the evaluation-time attack suite used in this thesis---namely additive noise, speed/resampling perturbation, low-pass filtering, crop-and-pad edits, bit-depth reduction, and MP3/AAC round-trips\,[9][10]?
  \item \textbf{RQ3 (Localized, low-latency detection):} On the same LibriSpeech benchmark, can the detector localize watermark presence and recover the 4-bit payload from short prefixes (0.25/0.5/1.0/2.0~s) while maintaining a low false-positive rate?
  \item \textbf{RQ4 (Integration status, only partially addressed):} What do the obtained watermark-prefix results imply for a future integrated ``watermark + user-side detector'' middleware, and which end-to-end latency components remain unmeasured in the current prototype?
\end{itemize}

\section{Proposed Approach (Overview)}
At the \emph{system-design} level, the proposed safety layer has two complementary branches:
\begin{enumerate}[leftmargin=1.25em]
  \item \textbf{Assistant-output watermarking for provenance (\emph{implemented in this thesis}).} An embedded, machine-detectable watermark is added to assistant speech. Any audio claiming to originate from the assistant can then be checked quickly by a localized decoder; forged or tampered outputs that lack the expected signature can be rejected.
  \item \textbf{User-side deepfake/replay detection for input integrity (\emph{conceptual design / future extension}).} Incoming user audio would be screened by a robust detector to reduce the risk of spoofed commands, replay attacks, or cloned-voice injections.
\end{enumerate}
Thus, the overall middleware vision is dual-branch, but the current thesis \textbf{experimentally validates only the first branch}. The second branch is discussed to motivate future integration, not to claim an implemented end-to-end system.

\section{Scope, Assumptions, and Threat Model}
\textbf{Implemented scope.} The empirical work in this thesis is limited to watermarking of assistant-originated speech on 16~kHz LibriSpeech audio and evaluation under a fixed black-box attack suite. The user-side deepfake/replay detector, full V2V deployment, and full pipeline latency measurement are outside the current implemented scope.

\textbf{Adversary capabilities.} We assume attackers can: (i) synthesize convincing target voices using modern TTS/VC; (ii) replay high-quality recordings; (iii) apply common signal transforms (compression, resampling, additive noise, filtering, editing) and attempt naive watermark removal. \textbf{Out-of-scope.} We do not consider physical compromise of the user device, sensor hardware modification, or attackers with perfect knowledge of and white-box access to the exact watermarking key and embedding internals.

\section{Evaluation Criteria and Datasets (Overview)}
In the \emph{implemented} scope, evaluation focuses on watermark-specific metrics: detection quality (TPR/FPR), payload recovery (bit and message accuracy), localization quality, and short-prefix verification behavior. The experiments use \textbf{LibriSpeech only}, and robustness is assessed under the implemented attack suite: additive noise, speed/resampling perturbation, low-pass filtering, crop-and-pad, bit-depth reduction, and MP3/AAC compression. Qualitative audio inspection is included, whereas formal MOS/A--B listening studies and the user-side detector evaluation remain future work. Prefix-level latency is reported, but \textbf{full end-to-end V2V middleware latency is only partially addressed} in the current prototype.

\section{Contributions}
\begin{itemize}[leftmargin=1.25em]
  \item \textbf{[Implemented in this thesis]} \textbf{Localized assistant-output watermarking on LibriSpeech.} A curriculum-trained waveform-domain watermarking pipeline for assistant speech that learns a stable 4-bit watermark and supports localized watermark-presence estimation.
  \item \textbf{[Implemented in this thesis]} \textbf{Low-latency prefix-aware verification.} A detector and training strategy that improve payload recovery from short audio prefixes, making watermark verification more compatible with streaming conversational settings.
  \item \textbf{[Implemented in this thesis]} \textbf{Reproducible robustness benchmark for the current prototype.} An evaluation protocol on LibriSpeech covering additive noise, speed/resampling, low-pass filtering, crop-and-pad, bit-depth reduction, and MP3/AAC round-trips, with reporting of TPR/FPR and payload-recovery metrics.
  \item \textbf{[Implemented in this thesis]} \textbf{Prototype codebase focused on the watermarking component.} An open-source experimental pipeline comprising staged training, evaluation, robustness sweeps, and qualitative rendering scripts for the assistant-output watermarking branch.
  \item \textbf{[Conceptual design / future extension]} \textbf{Integration path to a full safety middleware.} The thesis outlines how the implemented watermarking branch could later be combined with a user-side deepfake/replay detector into a bilateral ``watermark + detector'' middleware, but that second branch is not yet implemented here.
\end{itemize}
